---
title: Motivations and Barriers to Vaccination during Covid-19
author:
  - name: Megan Richard
    affil: 1
    orcid: '0000-0000-0000-0000'
affiliation:
  - num: 1
    address: Brown School, Washington University in St. Louis
column_numbers: 4
output: 
  posterdown::posterdown_html:
    self_contained: true
logoright_name: avatar-2alphas.png
logoleft_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/betterhexlogo.png
poster_height: "36in"
poster_width: "48in"
body_textsize: "36px"
sectitle_textsize: "48px"
primary_colour: "#a51417"
bibliography: packages.bib
knit: pagedown::chrome_print
---

```{r setup, include=FALSE}
#| echo: false
#| results: false
#| warning: false
#| error: false
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  out.width = "90%",
  out.height = "110%",
  dpi = 400)
```

```{css, echo=FALSE}
table.Rtable1 {
    font-family: "Times New Roman", serif;
    border-collapse: collapse;
    font-size: 28pt;
    margin-left: auto;
    margin-right: auto
}
```

# Research Purpose

-   This research seeks to answer the following question: Are age, political beliefs, and prior flu shot associated with the intention to get the COVID-19 vaccine for adults in the United States?

# Data Description

-   Data were collected in October 2020 through a Google Survey disseminated on social media

-   Survey questions were chosen by evaluating existing literature [@rasul2023; @ruiz2021; @albrecht2022]

-   After removing observations missing data on the 4 relevant variables, 4 variables and 511 observations will be examined for this project

-   The codebook can be downloaded from [this link](https://github.com/OxCGRT/covid-policy-tracker/blob/master/documentation/codebook.md)

# Variable Measurement

Outcome:

-   Intention to get the COVID-19 vaccine (categorical): Yes, No, Don't Know/Not Sure

Predictors:

-   Political beliefs (categorical): Very Liberal, Somewhat Liberal, Mixed, Somewhat Conservative, Very Conservative

-   Flu shot in the past 12 months (categorical); Yes, No

-   Age in years (continuous)

```{r include=FALSE , warning=FALSE, message=FALSE}

#Open Tidyverse 
library(package = "tidyverse") 
library(forcats)

# Import the data 
covid <- read.csv(file = "COVID-19 Vaccine Survey Responses.csv")

#take out international - does not align with research question
covid.dirty <- covid %>% 
  select( Once.it.is.available..do.you.intend.to.get.the.vaccine.for.COVID.19.,
          Have.you.received.a.flu.shot.in.the.past.12.months., 
          What.is.the.highest.degree.or.level.of.school.you.have.completed., 
          Which.most.closely.aligns.with.your.political.beliefs.,
          I.am.at.risk.for.contracting.COVID.19.,
          COVID.19.can.be.a.serious.disease.,
          The.COVID.19.vaccine.will.prevent.me.from.contracting.COVID.19., 
          There.are.many.risks.associated.with.getting.the.COVID.19.vaccine.,
          State.of.Residence,
          What.is.your.age.in.years.) %>% 
  mutate(State.of.Residence = recode_factor(.x = State.of.Residence,
                                        `International` = NA_character_)) %>% 
  filter(!is.na(State.of.Residence))

#Recode and label variables, ensure missing values are properly coded
covid.small <- covid.dirty %>% 
  select( Once.it.is.available..do.you.intend.to.get.the.vaccine.for.COVID.19.,
          Have.you.received.a.flu.shot.in.the.past.12.months., 
          Which.most.closely.aligns.with.your.political.beliefs.,
          What.is.your.age.in.years.) %>% 
  #mutate_all(as.factor) %>% 
  rename(vaccine.intention = Once.it.is.available..do.you.intend.to.get.the.vaccine.for.COVID.19.) %>% 
  rename(flu.shot.history = Have.you.received.a.flu.shot.in.the.past.12.months.) %>%
  mutate(flu.shot.history = recode_factor(.x = flu.shot.history,
                                          `Don't know/Not sure` = NA_character_)) %>% 
  rename(political.beliefs = Which.most.closely.aligns.with.your.political.beliefs.) %>% 
  mutate(political.beliefs = recode_factor(.x = political.beliefs,
                                         '.' = "Prefer not to answer")) %>% 
  rename(age.years = What.is.your.age.in.years.) %>% 
  mutate(age.years = as.numeric(as.character(age.years))) %>% 
  drop_na()


#summarize the cleaned data 
summary(covid.small)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}

#Creating levels for each variable so they are ordered in logical sense on graphs below
covid.small$flu.shot.history <- factor(covid.small$flu.shot.history,
                                       levels = c("Yes", "No"))
covid.small$vaccine.intention <- factor(covid.small$vaccine.intention,
                                        levels = c("Yes", "No", "Don't Know/Not sure"))
covid.small$political.beliefs <- factor(covid.small$political.beliefs, 
                                        levels = c("Very Conservative", "Somewhat Conservative", "Mixed", "Somewhat Liberal", "Very Liberal", "Prefer not to answer"))

```

```{r include = FALSE}
# check age distribution
hist(covid.small$age.years)
```

# Participant Characteristics

```{r echo=FALSE, warning=FALSE, message=FALSE}

#Open TableOne package 
library(package = "table1")
library(package = "tableone")

#Creating table to show descriptive statistics in percents 
#add labels to the table
labelled::var_label(x = covid.small) <- c("Intends to get the vaccine for COVID-19",
                                          "Flu shot in the past 12 months",
                                          "Political beliefs",
                                          "Age in years (median, IQR)")

# use print to show table with labels and percent
table1(~ vaccine.intention + flu.shot.history +
         political.beliefs + age.years,
       data = covid.small,
       render.continuous = "Median, IQR",
       caption = "Characteristics of 511 Participants in a Covid-19 Vaccination Survey, 2022.")


```

-   Nearly half of participants intended to get the covid vaccine (47.5%; n = 242)

-   Most survey participants had received a flu shot in the past year (64.6%; n = 330)

-   Most participants were somewhat or very liberal (55.3%; n = 282)

-   The median age of participants was 26 years old (IQR = 25)

```{r include = FALSE}
# hide this but use info for bullet points below
table1(~ flu.shot.history + 
         political.beliefs + age.years | vaccine.intention,
       data = covid.small,
       render.continuous = "Median, (Q1, Q3)")
```

# Vaccine Intention and Flu Shot History

-   Figure 1 shows a higher percentage of people who intend to get the COVID-19 vaccine have had the flu shot and the opposite is true for those not intending to get the COVID-19 vaccine

    -   Of those who intend to get the COVID-19 vaccine, 79.3% of them had a flu shot in the last year and 20.7% did not have a flu shot

    -   Of those who do not intend to get the COVID-19 vaccine, only 37.5% of them had a recent flu shot and 62.5% did not have a flu shot

```{r echo=FALSE, message=FALSE, warning=FALSE}
covid.small <- covid.small %>% 
  mutate(vaccine.intention = recode_factor(vaccine.intention,
    "Don't Know/Not sure" = "Unsure Intent to get Covid Vaccine",
    "No" = "No Intent to get Covid Vaccine (No)",
    "Yes" = "Intent to get Covid Vaccine (Yes)",
  ))

#Figure 1 - Flu Shot History Graphed with Percent in Intention to get COVID-19 Vaccine
figure1 <- covid.small %>% 
  drop_na(vaccine.intention, flu.shot.history) %>% 
  group_by(vaccine.intention, flu.shot.history) %>% 
  count() %>% 
  group_by(vaccine.intention) %>% 
  mutate(percent = 100*(n/sum(n))) %>% 
  ggplot(aes(x = flu.shot.history, fill = vaccine.intention,
             y = percent)) +
  geom_col(position = "dodge") +
  theme_bw(base_size = 18, base_family = 'serif') + 
  theme(strip.background = element_rect(fill = "#eff3ff")) +
  labs(x = "Have you received a flu shot\nin the past 12 months?", 
       y = "Percent in vaccine intention group",
       title = "Figure 1: Vaccine Intention & Flu Shot History\n(2022; n = 511)") +
  scale_fill_manual(values = c("#1b9e77", "#d95f02", "#7570b3"), 
                    guide = "none") +
  facet_wrap(facets = 'vaccine.intention', nrow = 3) +
  coord_flip()
figure1
```

```{r include = FALSE}
descr::CrossTable(x = covid.small$flu.shot.history,
                  y = covid.small$vaccine.intention,
                  sresid = TRUE,
                  expected = TRUE,
                  prop.r = FALSE,
                  prop.c = FALSE,
                  prop.t = FALSE,
                  chisq = TRUE)
# assumptions met, expected values are large in all cells
```

-   Flu shot history and vaccine intention are both categorical so I used chi-squared to determine whether they were statistically significantly associated

    -   All chi-squared assumptions were met including independent observations, both variables were nominal or ordinal, and more than 80% of expected values were greater than 5

-   There is a statistically significant association between vaccine intention and flu shot history ($\chi^2$ = 56.35; p \< .05)

-   Standardized residuals showed which groups were driving the statistically significant result

    -   Significantly **fewer than expected** people who had a flu shot did not intend to take the COVID-19 vaccine (std res = -3.30) and **more than expected** who had a flu shot intended to take the vaccine (std res = 2.87)

    -   Significantly **more than expected** people with no flu shot did not intend to take the COVID-19 vaccine (std res = 4.44) and **fewer than expected** with no flu shot intended to take the vaccine (std res = -3.87)

# Vaccine Intention and Political Beliefs

-   Figure 2 suggests that political beliefs may be associated with intention to get the COVID-19 vaccine

    -   Of those who intend to get the COVID-19 vaccine, 50% were very liberal and 24.8% were somewhat liberal

    -   Of those who do not intend to get the COVID-19 vaccine, 32.3% were politically mixed and 21.9% were somewhat conservative

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=7}
#Figure 3 - Political Beliefs Graphed with Percent in Intention to get COVID-19 Vaccine
figure3 <- covid.small %>% 
  drop_na(vaccine.intention, political.beliefs) %>% 
  group_by(vaccine.intention, political.beliefs) %>% 
  count() %>% 
  group_by(vaccine.intention) %>% 
  mutate(percent = 100*(n/sum(n))) %>%
  ggplot(aes(x = political.beliefs, fill = vaccine.intention,
             y = percent)) +
  geom_col(position = "dodge") +
  coord_flip() +
  theme_bw(base_size = 18, base_family = 'serif') + 
  theme(strip.background = element_rect(fill = "#eff3ff")) + 
  labs(x = "Political beliefs", 
       y = "Percent in vaccine intention group",
       title = "Figure 2: Vaccine Intention &\nPolitical Beliefs (2022; n = 511)") +
  scale_fill_manual(values = c("#1b9e77", "#d95f02", "#7570b3"),
                    guide = "none") +
  facet_wrap(facets = 'vaccine.intention', nrow = 3)
figure3
```

```{r include = FALSE}
descr::CrossTable(x = covid.small$political.beliefs,
                  y = covid.small$vaccine.intention,
                  sresid = TRUE,
                  expected = TRUE,
                  prop.r = FALSE,
                  prop.c = FALSE,
                  prop.t = FALSE,
                  chisq = TRUE)
# assumptions are met, expected values are large in all cells
```

-   Political beliefs and vaccine intention are both categorical so I used chi-squared to determine whether they were statistically significantly associated

    -   All chi-squared assumptions were met including independent observations, both variables were nominal or ordinal, and more than 80% of expected values are greater than 5

-   I found a statistically significant association between vaccine intention and political beliefs ($\chi^2$ = 142.55; p \< .05).

-   Standardized residuals showed which groups were driving the statistically significant result

    -   Significantly **more very liberal people than expected were intending to take the COVID-19 vaccine** (std res = 4.83) while **significantly fewer than expected** who preferred not to answer, were politically mixed, or were somewhat conservative were intending to take the vaccine (std res \< -2)

    -   Significantly **more than expected** people who preferred not to answer, were politically mixed, or were very conservative were not intending to take the COVID-19 vaccine (std res \> 2) while **significantly fewer than expected** who were very liberal or somewhat liberal were intending to not take the vaccine (std res \< -2)

# Vaccine Intention and Age

-   The median age is highest in the "No" vaccine intention group (median = 39.5, IQR = 23.8--50.0) and lowest in the "Yes" vaccine intention group (median = 24.0, IQR = 22.0--41.8)

-   The large difference suggests that age is likely associated with vaccine intention and that there are more older participants who do not intend to get the COVID-19 vaccine

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(gghalves)
figure4 <- covid.small %>% 
  ggplot(aes(fill = fct_rev(vaccine.intention),
             y = age.years,
             x = vaccine.intention)) +
  geom_half_violin(aes(fill = fct_rev(vaccine.intention)), alpha = .4, side = "r") +
  geom_half_point(aes(color = fct_rev(vaccine.intention)), alpha = .8, side = "l") +
  theme_bw(base_size = 18, base_family = 'serif') + 
  labs(y = "Age in Years", 
       x = "Intent to Get Covid Vaccine",
       title = "Figure 3: Vaccine Intention & Age in Years\n(2022; n = 511)") +
  scale_fill_manual(values = c("#1b9e77", "#d95f02", "#7570b3"),
                    guide = "none") +
  scale_color_manual(values = c("#1b9e77", "#d95f02", "#7570b3"),
                    guide = "none") +
  scale_x_discrete(
    labels = function(x) str_wrap(x, width = 15),
    drop = FALSE
  ) +
  coord_flip()
  
figure4
```

```{r include = FALSE}
# fails the assumption of normal distribution within groups, the Yes group is very right skewed
# the don't know group is somewhat right skewed
# Kruskal-Wallis test
kruskal.test(formula = age.years ~ vaccine.intention,
             data = covid.small)
# stat sig so do follow-up test
dunn.test::dunn.test(x = covid.small$age.years,
                     g = covid.small$vaccine.intention,
                     method = "bonferroni")

```

-   The ANOVA assumption of normal distribution within groups was not met with the distribution of age being very right skewed in the group who intended to get the COVID-19 vaccine

-   Age is continuous and vaccine intention is categorical with 3 categories so I used the Kruskal-Wallis test to test whether the distribution of ranked age was the same across vaccine intention groups

-   There was a statistically significant difference in age distribution across the vaccine intention groups (KW $\chi^2$ = 11.63; p = .003)

-   A post-hoc Dunn's test can determine which groups are statistically significantly different from each other

    -   The mean age of the group with no intention to vaccinate for COVID-19 was statistically significantly higher (diff = 3.35 years) compared to the group that intended to vaccinate for COVID-19 (p = .001)

# Summary and Recommendations

Age, having a recent flu vaccine, and political beliefs were all statistically significantly associated with intent to get a COVID-19 vaccination. Consistent with prior studies, I found that people who had a flu vaccination in the last year, had more liberal political beliefs, and were younger were more likely to intend to get the COVID-19 vaccine [@rasul2023; @ruiz2021; @albrecht2022].

Increasing the uptake of flu vaccination and developing health communication campaigns that reach older and more conservative audiences could both increase the uptake of COVID-19 vaccination.

# References

::: {#refs}
:::
